---
title: "Task 2: Lizard Length to Weight NLS"
author: "Claire Meuter"
date: "2023-02-12"
output: 
  html_document:
   code_folding: hide
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(broom)
library(here)
library(Metrics)
library(kableExtra)

```

```{r}
#read in the data
lizards <- read_csv(here("lizards.csv"))
```

This rmd wrangles, explores, and models data from lizard traps in the Jornada Basin LTER site. 

Source: Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f

### 1. Select model
Fit a snout length to weight model of the following form to all lizards in your dataframe.
\begin{equation}
W=a(SVL)^b
\end{equation}

Let's create a function in R that represents this equation
```{r}
liz_func <- function(a, b, svl){
  weight <- a*(svl^b)
  return(weight)
}
```

### 1a. Making an intial guess
Weight is given by W, snout to vent length by SVL, and a and b are the parameters that need to be fitted. 
```{r}
# OLS regression on log-transformed model
# Since we know the model is exponential in nature, we could log transform the data. If we do a standard OLS regression on the log transformed data, we can get approximations of the parameters from the regression coefficients
 
guess_model <- lm(log(lizards$weight) ~ log(lizards$SV_length), data = lizards)


#c. Using the coefficients function, we can then supply the NLS start list with the regression coefficients. Hint: Because you log transformed the data, you will have to mathematically transform the intercept coefficient to get the guess for parameter a.

#Now I can look at what my coefficients look like for the guess model: 
#coefficients(guess_model)

#assign guesses for parameters a and b 
a <- exp((guess_model$coefficients[1]))
b <- guess_model$coefficients[2]
```

### 2. Run NLS 
```{r}
#run nls 
lizards_nls <- nls(weight ~ liz_func(a, b, SV_length),
                   data = lizards, 
                   start = list(a = a, b = b))
#return results in tidy table 
lizards_nls_tidy <- tidy(lizards_nls)

lizards_nls_tidy %>% 
  kable(caption= "Table 1: NLS results",
        col.names = c("Parameter", 
                      "Coefficient", 
                      "Standard Error", 
                      "Statistic", 
                      "P-Value")) %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = "striped",
                position = "left")
  
  
```

Now I can predict weight based off this model, and display the results in plot. 
```{r}
#run predictions of weight 
lizards_pred <- lizards %>% 
  mutate(predict = predict(lizards_nls, newdata=.))

#plot how well weight prediction matches actual weight 
ggplot(data = lizards_pred) +
  #include true weight pts 
  geom_point(aes(x = SV_length, y = weight, color = sex)) +
  scale_color_discrete(labels = c("Female", "Male")) +
  #add our prediction data
  geom_line(aes(x = SV_length, y = predict),
            color = "black",
            size = 1) +
  labs(x = "Snout to vent length (mm)",
       y = "Weight (g)") +
  theme_minimal()

```

**Figure 1.** Snout to vent lengths (mm) and body weights (g) for male and female lizards. Female observations are in red, and male observations in blue. The observation points are overlaid with a line representing the fit of the NLS predicted model. 

### 3. Filtering for Male Western Whiptail lizard (*C. tigrisatus*)
```{r}
#creating new data subset from lizards data 
lizards_c <- lizards %>% 
  filter(spp == "CNTI", #data code for C. tigrisatus
         sex == "M")

# now we can recreate what we did above, but only with C. tigrisatus 
# intial guess model 
guess_model_c <- lm(log(lizards_c$weight) ~ log(lizards_c$SV_length), data = lizards_c)

#explore my OLS coefficients 
coefficients(guess_model_c)

#assign guesses for parameters a and b 
a_c <- exp((guess_model_c$coefficients[1]))
b_c <- guess_model_c$coefficients[2]

# run NLS 
lizards_nls_c <- nls(weight ~ liz_func(a, b, SV_length),
                   data = lizards_c, 
                   start = list(a = a_c, b = b_c))
#return results in tidy table 
lizards_nls_c_tidy <- tidy(lizards_nls_c)

lizards_nls_c_tidy %>% 
  kable(caption= "Table 1: NLS results",
        col.names = c("Parameter", 
                      "Coefficient", 
                      "Standard Error", 
                      "Statistic", 
                      "P-Value")) %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = "striped",
                position = "left")

```
```{r}
#predict the weight based on model
#run predictions of weight 
lizards_pred_c <- lizards_c %>% 
  mutate(predict = predict(lizards_nls_c, newdata=.))


#getting RMSE for predictions 
lizards_rmse <- data.frame(rmse_c = rmse(lizards_pred_c$weight, 
                                            lizards_pred_c$predict),
                           rmse_cnti_v_gen = rmse(lizards_pred_c$weight, 
                                                  lizards_pred$predict),
                           rmse_gen = rmse(lizards_pred$weight, 
                                           lizards_pred$predict))

#plot how well weight prediction (species specific and general model)matches actual weight 
ggplot(data = lizards_pred_c) +
  #include true weight pts 
  geom_point(aes(x = SV_length, y = weight), color = "black") +
  #add our prediction data for species specific model
  geom_line(aes(x = SV_length, y = predict),
            color = "darkorange",
            size = .8)  +
  #line for general nls model 
  geom_line(data = lizards_pred,
            aes(x = SV_length,
                y = predict),
            color = "blue") +
  labs(x = "Snout to vent length (mm)",
       y = "Weight (g)") +
  theme_minimal()

```

**Figure 1.** Snout to vent lengths (mm) and body weights (g) for male *C. tigrisatus*. The orange line displays the fit of the  *C. tigrisatus* species-specific NLS model, which has an RMSE of `r round(lizards_rmse[1],3)` g.The blue line represents the general NLS model for all species of lizards, which has an RMSE of `r round(lizards_rmse[3],3)` g. Because the general model has a lower RMSE, we would want to use that model.
